---
title: "Island Vibes"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
# load packages
if (!require(tidyverse)) {
  install.packages("tidyverse")
}
if (!require(scales)) {
  install.packages("scales")
}
if (!require(patchwork)) {
  install.packages("patchwork")
}
if (!require(jcolors)) {
  install.packages("jcolors")
}
if (!require(knitr)) {
  install.packages("knitr")
}
if (!require(lubridate)) {
  install.packages("lubridate")
}
if (!require(tidymodels)) {
  install.packages("tidymodels")
}
if (!require(janitor)) {
  install.packages("janitor")
}
if (!require(kableExtra)) {
  install.packages("kableExtra")
}
if (!require(reticulate)) {
  install.packages("reticulate")
}
if (!require(fuzzyjoin)) {
  install.packages("fuzzyjoin")
}
if (!require(mgcv)) {
  install.packages("mgcv")
}
if (!require(zoo)) {
  install.packages("zoo")
}
if (!require(ggrepel)) {
  install.packages("ggrepel")
}
if (!require(spotifyr)) {
  install.packages("spotifyr")
}
if (!require(corrr)) {
  install.packages("corrr")
}
if (!require(tidytext)) {
  install.packages("tidytext")
}
if (!require(LiblineaR)) {
  install.packages("LiblineaR")
}
if (!require(keyring)) {
  install.packages("keyring")
}


# this loads custom plot theme from saved location

source('~/Documents/Projects/tools/theme_main.R')

# loading in helper functions

source('~/Documents/Projects/ExploreF1/helper_functions_track_evo.R')

```

  
```{r, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}

theme_main_add <- function() {
  
    theme(plot.title = element_text(hjust = 0),
        plot.subtitle = element_text(face = "italic",
                                     hjust = 0),
        plot.background = element_rect(fill = NA,
                                       color = "black",
                                       size = 2)
        )
  
}

```

```{r}

id <- "60714a4297514197b37355b07367cb0a"

Sys.setenv(SPOTIFY_CLIENT_ID = id)
Sys.setenv(SPOTIFY_CLIENT_SECRET = key_get(service = "secret", username = "colin_renville_secret"))
access_token <- get_spotify_access_token()

```


```{r}

my_id <- "12121211477"
my_plists <- get_user_playlists(my_id)

island_vibes_df <- my_plists %>%
  dplyr::filter(stringr::str_detect(name, "Island Vibes"),
                id == "4B2lrMRvAnzWYeFGchNEGC") %>%
  unnest(images) 

```


```{r, echo=FALSE, eval=FALSE}


tracks1 <- get_playlist_tracks(island_vibes_df$id)
tracks2 <- get_playlist_tracks(island_vibes_df$id, offset = 100)
tracks3 <- get_playlist_tracks(island_vibes_df$id, offset = 200)
tracks <- rbind(tracks1, tracks2, tracks3)

features1 <- get_track_audio_features(tracks1$track.id)
features2 <- get_track_audio_features(tracks2$track.id)
features3 <- get_track_audio_features(tracks3$track.id)
features <- rbind(features1, features2
                  # , features3
                  )

track_features2 <- tracks %>%
  left_join(features, by = c("track.id" = "id")) %>%
  select("track.id", "track.artists", "track.name", "track.duration_ms", "track.popularity",
         danceability,
         energy,
         key,
         loudness,
         mode,
         speechiness,
         acousticness,
         instrumentalness,
         liveness,
         valence,
         tempo
         # type
         ) %>%
  unnest("track.artists",
         names_repair = "unique") %>%
  group_by(track.id) %>%
  mutate(song_rownum = row_number() # , 
         # artists_name = paste0(unlist(name), collapse = ","),
         # id_name = paste0(unlist(id), collapse = ",")
         ) %>%
  ungroup() %>%
  filter(song_rownum == "1") %>% # only selecting row for primary artist of a song
  select(-c(href, type, uri, external_urls.spotify, song_rownum))


```


```{r}

# scale the above chunk into a function based on island_vibes_df$tracks.total

create_playlist_song_features <- function(playlist_id) {
  
  # debug
  # playlist_id <- "4B2lrMRvAnzWYeFGchNEGC"
  # playlist_track_ct <- 471
  
  
  if (exists("my_plists")) {
    
    assertthat::assert_that(inherits(my_plists, "data.frame"),
                            msg = "my_plists must be a dataframe.")
    
  } else {
      
    print("my_plists object does not exist!")
    
  }
  
  playlist_df <- my_plists %>%
    dplyr::filter(id == playlist_id) %>%
    select(id, name, tracks.total)
  
  # pull track count
  ## need to ensure that there is only a single row / playlist ? 
  
  playlist_track_ct <- playlist_df %>%
    dplyr::pull(as.numeric(tracks.total))
  
  # derive the offsets needed, given a playlist's playlist_track_ct
  
  playlist_track_max_offset <- plyr::round_any(playlist_track_ct, 100, f = floor)
  
  # if between 100 and 199, then offset == 100
  # if between 200 and 299, then offset == 200, .etc
  
  playlist_offset <- seq(0,playlist_track_max_offset, by = 100)
  
  playlist_tracks_df <- playlist_offset %>%
    map_df(~ get_playlist_tracks(playlist_df$id,
                                 offset = .x) %>%
             mutate(offset_group = as.factor(as.character(.x)))) 
  
  # 100 row limit still applies for get_track_audio_features(), with no offset argument
  ## batch applying get_track_audio_features in groups of 100
  
  playlist_features_df <- playlist_offset %>%
    map_df(~ get_track_audio_features(ids = playlist_tracks_df$track.id[playlist_tracks_df$offset_group == .x]))
  
  # join song and feature dataframes, select relevant columns
  
  playlist_tracks_features_df <- playlist_tracks_df %>%
    left_join(playlist_features_df, by = c("track.id" = "id")) %>%
    select("track.id", "track.artists", "track.name", "track.duration_ms", "track.popularity",
           danceability,
           energy,
           key,
           loudness,
           mode,
           speechiness,
           acousticness,
           instrumentalness,
           liveness,
           valence,
           tempo
           # type
           ) %>%
    unnest("track.artists",
           names_repair = "unique") %>%
    group_by(track.id) %>%
    mutate(song_rownum = row_number() # , 
           # artists_name = paste0(unlist(name), collapse = ","),
           # id_name = paste0(unlist(id), collapse = ",")
           ) %>%
    ungroup() %>%
    filter(song_rownum == "1") %>% # only selecting row for primary artist of a song
    select(-c(href, type, uri, external_urls.spotify, song_rownum))
  
  return(playlist_tracks_features_df)
  
}


track_features <- create_playlist_song_features(playlist_id = "4B2lrMRvAnzWYeFGchNEGC")

# min(track_features2 == track_features)

```


### Analyze Island Vibes

```{r}

track_features %>%
  select(track.popularity:tempo) %>%
  na.omit() %>%
  correlate() %>%
  rearrange() %>%
  network_plot(colours = c("orange", "white", "midnightblue"))

```


```{r}

# spotify API only allows 50 IDs at a time

island_vibes_artist_ids <- track_features$id

track_feature_artists1 <- paste0(island_vibes_artist_ids[1:50], collapse=",")
track_feature_artists2 <- paste0(island_vibes_artist_ids[51:100], collapse=",")
track_feature_artists3 <- paste0(island_vibes_artist_ids[101:150], collapse=",")
track_feature_artists4 <- paste0(island_vibes_artist_ids[151:200], collapse=",")
track_feature_artists5 <- paste0(island_vibes_artist_ids[201], collapse=",")

artists_genre1 <- spotifyr::get_artists(ids = as.character(track_feature_artists1))
artists_genre2 <- spotifyr::get_artists(ids = as.character(track_feature_artists2))
artists_genre3 <- spotifyr::get_artists(ids = as.character(track_feature_artists3))
artists_genre4 <- spotifyr::get_artists(ids = as.character(track_feature_artists4))
artists_genre5 <- spotifyr::get_artists(ids = as.character(track_feature_artists5))


track_feature_artists <- rbind(artists_genre1,
                               artists_genre2,
                               artists_genre3,
                               artists_genre4,
                               artists_genre5)


artists_genre_dis <- track_feature_artists %>%
  select(id, name, genres, popularity, followers.total) %>%
  distinct() %>%
  group_by(id) %>%
  mutate(genres2 = paste0(unlist(genres), collapse = ", ")) %>%
  ungroup() %>%
  select(-c(name, genres))

track_features <- track_features %>%
  left_join(artists_genre_dis, by = c("id" = "id"))

tidy_artists <- track_features %>%
  select(track.id, track.name, name, genres2, track.popularity) %>%
  unnest_tokens(word, genres2) %>%
  distinct()

tidy_artists %>% head(15)

```


```{r}

# scale the above chunk into a function to wrangle artists from playlist

create_playlist_artist_features <- function(playlist_id) {
  
  # debug
  # playlist_id <- "4B2lrMRvAnzWYeFGchNEGC"
  # playlist_track_ct <- 471
  
  
  if (exists("track_features")) {
    
    assertthat::assert_that(inherits(track_features, "data.frame"),
                            msg = "track_features must be a dataframe.")
    
  } else {
      
    print("track_features object does not exist!")
    
  }
  
  track_features_track_ct <- nrow(track_features)
  
  track_features_max_offset <- plyr::round_any(track_features_track_ct, 100, f = ceiling) + 1
  
  track_features_offset <- seq(1, track_features_max_offset, by = 50)
  
  # ...
  
  }


```



```{r}

tidy_artists %>%
  group_by(word) %>%
  summarise(
    n = n(),
    track.popularity = mean(track.popularity)
  ) %>%
  ggplot(aes(n, track.popularity)) +
  geom_hline(
    yintercept = mean(tidy_artists$track.popularity), lty = 2,
    color = "gray50", size = 1.5
  ) +
  geom_jitter(color = "midnightblue", alpha = 0.7) +
  geom_text_repel(aes(label = word),
                  max.overlaps = 20) +
  labs() +
  scale_x_log10() +
  theme_main() +
  theme_main_add()

```


```{r}

set.seed(123)
iv_split <- initial_split(track_features, strata = track.popularity)
iv_train <- training(iv_split)
iv_test <- testing(iv_split)

set.seed(234)
iv_folds <- vfold_cv(iv_train, strata = track.popularity, v = 100)
iv_folds

```


```{r}

require("devtools")
# install_github("tidymodels/textrecipes")

library(textrecipes)

iv_rec <- recipe(track.popularity ~ genres2, data = iv_train) %>%
  step_tokenize(genres2) %>%
  step_tokenfilter(genres2, max_tokens = 100) %>%
  step_tfidf(genres2)

## just to check this works
prep(iv_rec) %>% bake(new_data = NULL)

```

### Model: Popularity ~ genre text features

```{r}

svm_spec <-
  svm_linear() %>%
  set_mode("regression")

svm_spec

```


```{r}

svm_wf <- workflow(iv_rec, svm_spec)

```


```{r}

# install.packages("doParallel")
library(doParallel)
doParallel::registerDoParallel()
contrl_preds <- control_resamples(save_pred = TRUE)

svm_rs <- fit_resamples(
  svm_wf,
  resamples = iv_folds,
  control = contrl_preds
)

collect_metrics(svm_rs)

```


```{r}

collect_predictions(svm_rs) %>%
    mutate(mod = "SVM") %>%
  ggplot(aes(track.popularity, .pred)) +
  geom_abline(lty = 2, color = "gray50", size = 1.2) +
  geom_jitter(aes(color = id),
              width = 0.5, alpha = 0.5,
              show.legend = FALSE) +
  facet_wrap(vars(mod)) +
  coord_fixed() +
  theme_main() +
  theme_main_add()

```


```{r}

final_fitted <- last_fit(svm_wf, iv_split)
collect_metrics(final_fitted) ## metrics evaluated on the *testing* data

final_wf <- extract_workflow(final_fitted)


```

```{r}
  
final_wf %>%
 recipes::tidy() %>%
  filter(term != "Bias") %>%
  group_by(estimate > 0) %>%
  slice_max(abs(estimate), n = 10) %>%
  ungroup() %>%
  mutate(term = str_remove(term, "tfidf_most_memorable_characteristics_")) %>%
  ggplot(aes(estimate, fct_reorder(term, estimate), fill = estimate > 0)) +
  geom_col(alpha = 0.8) +
  scale_fill_discrete(labels = c("low ratings", "high ratings")) +
  labs(y = NULL, fill = "More from...") +
  theme_main() +
  theme_main_add()

```


### EDA: What are the characteristics of Island Vibes songs?

```{r}

# converting song features from wide to long

track_features_long <- track_features %>%
  select(track.id, id, name, track.name, danceability:tempo) %>%
  pivot_longer(cols = danceability:tempo,
               names_to = "song_feature",
               values_to = "song_feature_value")

track_features %>% head(1)
track_features_long %>% head(11) 

```


```{r}

track_features_long %>%
  ggplot(aes(x = song_feature_value,
             fill = song_feature)) +
  geom_histogram(position = position_dodge(),
                 # bins = 2000,
                 color = "black",
                 show.legend = FALSE) +
  labs(title = "Island Vibes: Song feature distributions",
       subtitle = "Songs we will be looking for will be danceable and generally have high valence and loudness and low instrumentalness, acousticness and speechiness", 
       x = "Feature Value",
       y = "Ct. songs") +
  # coord_cartesian(xlim = c(0,1)) +
  facet_wrap(~song_feature,
             scales = "free") +
  theme_main() +
  theme_main_add()

```


### Create Principal Components to combine variance of song features

```{r}

track_features_pca <- track_features %>% 
  select(track.id, track.name, id, name, danceability:tempo)

set.seed(123)

iv_recipe_pca <- recipe(~ ., data = track_features_pca) %>%
  update_role(track.id, track.name, id, name,
              new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

iv_pca_prep <- prep(iv_recipe_pca)

```

```{r}

tidy_pca <- tidy(iv_pca_prep, 2)

new_order <- tidy_pca %>% 
  group_by(component) %>% 
  do(data_frame(al=levels(reorder(interaction(.$component, .$terms, drop=TRUE), .$value)))) %>% 
  pull(al)

tidy_pca %>%
  mutate(component = fct_inorder(component)) %>%
  mutate(al=factor(interaction(component, terms), levels=new_order)) %>%
  ggplot(aes(value, al, fill = al)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component,
             scales = "free_y") +
  labs(y = NULL) +
  theme_main() +
  theme_main_add()

```

How much variance in song features is explained by each Principal Component?

```{r}

sdev <- iv_pca_prep$steps[[2]]$res$sdev
percent_variation <- sdev^2 / sum(sdev^2)

tibble(
  component = unique(tidy_pca$component),
  percent_var = percent_variation ## use cumsum() to find cumulative, if you prefer
) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(component, percent_var, group = 1)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = NULL, y = "Percent variance explained by each PCA component") +
  theme_main() +
  theme_main_add()


tibble(
  component = unique(tidy_pca$component),
  percent_var = cumsum(percent_variation) 
) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(component, percent_var, group = 1)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = NULL, y = "Percent variance explained by each PCA component") +
  coord_cartesian(ylim = c(0,1)) +
  theme_main() +
  theme_main_add()

```

```{r}

juice(iv_pca_prep) %>% View()

```

```{r}

# library(devtools)
# install_github("vqv/ggbiplot")
library(ggbiplot)


# juice(iv_pca_prep) %>%
#   ggbiplot()
  
  
  ggplot(aes(x = PC1, y = PC2, label = track.name)) +
  geom_point() +
  geom_text_repel(max.overlaps = 150,
                  size = 3) +
  geom_hline(yintercept = 0,
             color = "red",
             linetype = "dashed") +
  geom_vline(xintercept = 0,
             color = "red",
             linetype = "dashed") +
  theme_main() +
  theme_main_add()


```


PCA is not super great here, in that the variance seems too high for just 1-2 PCs to significantly explain.

### UMAP

```{r}

library(embed)


set.seed(123)

iv_recipe_umap <- recipe(~ ., data = track_features_pca) %>%
  update_role(track.id, track.name, id, name,
              new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors())

iv_umap_prep <- prep(iv_recipe_umap)

iv_umap_prep

```

```{r}

juiced_joined_umap <- juice(iv_umap_prep) %>%
  left_join(select(track_features, c("track.id",
                                     "valence",
                                     "acousticness",
                                     "energy",
                                     "loudness",
                                     "track.popularity")), by = c("track.id" = "track.id"))

juiced_joined_umap %>%
  ggplot(aes(UMAP1, UMAP2, color = track.popularity)) +
  geom_point(alpha = 0.7, size = 2) +
  # geom_text(check_overlap = TRUE, hjust = "inward", family = "IBMPlexSans") +
  labs(color = NULL) +
  theme_main() +
  theme_main_add()  

```




### Can we create a "bag" of songs to pick from and compare? 

- Directionally from frequent island vibes genres
  - probably reggae, country, fusion, rock, some rap

```{r}



```





